### 2013-05-07 doc/icml2013/evaluation.tex
Added bit of discussion, and minor changes

\section{Empirical Evaluation}
\label{sec:evaluation}

In the previous section, we showed that Spectral Experts is a consistent
algorithm that converges to the global solution at a $\frac{1}{\sqrt}$
rate. In this section, we will support these theoretical results with
experiments on simulated data. We will see that in the finite sample
regime, even when Spectral Experts does not converge to the optimal
answer, it provides robust initialization for local methods like EM.

\subsection{Experimental Setup}

@@

We solved the low rank recovery optimization routines for $M_2$ and
$M_3$ using \texttt{cvx}\citationneeded. 
%ainterior point method an interior proximal subgradient descent
%algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
%The algorithm was initialized with the regularized least squares
%solution and was run for 500 iterations or until convergence up to
%a tolerance of $10^{-3}$.
%Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$. 
As a baseline, we also implemented EM with two different
initializations. In the first, we used $\beta$s with random standard
Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
we initialized EM with the starting points given by Spectral Experts.

>>>>

\section{Empirical evaluation}
\label{sec:evaluation}

In the previous section, we showed that Spectral Experts is a consistent
algorithm that converges to the global solution at a $\frac{1}{\sqrt n}$
rate. In this section, we will support these theoretical results with
experiments on simulated data. We will see that in the finite sample
regime, even when Spectral Experts does not converge to the optimal
answer, it provides robust initialization for local methods like EM.

\subsection{Experimental setup}

@@

We solved the low rank recovery optimization routines for $M_2$ and
$M_3$ using an off-the-shelf solver for convex programs, CVX\cite{cvx}.
%ainterior point method an interior proximal subgradient descent
%algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
%The algorithm was initialized with the regularized least squares
%solution and was run for 500 iterations or until convergence up to
%a tolerance of $10^{-3}$.
%Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$. 
As a baseline, we also implemented EM with two different
initializations. In the first, we used $\beta$s with random standard
Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
we initialized EM with the starting points given by Spectral Experts.

### 2013-05-07 doc/icml2013/intro.tex
Added bit of discussion, and minor changes

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\Lop$ denote the operator norm (largest singular value),
let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
For a tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\Lop = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\Lop$
denote the average operator norm over all $p$ unfoldings.

For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
\Re^{D_p}, D_p = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)
= (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
In general, each component of $\cvec(X)$ is indexed by a vector of
counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
\}$ are the set of index vectors $k$ with that count profile.  
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

>>>>

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\op$ denote the operator norm (largest singular value),
let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
For a tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\op = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\op$
denote the average operator norm over all $p$ unfoldings.

For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
\Re^{C_{d,p}}, C_{d,p} = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)
= (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
In general, each component of $\cvec(X)$ is indexed by a vector of
counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
\}$ are the set of index vectors $k$ with that count profile.  
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-05-07 doc/icml2013/theory.tex
Added bit of discussion, and minor changes

We note that this requires more stringent positive definite constraints
when using features. For example, suppose $x = (1, t, t^2)$, the common
polynomial basis expansion, so that all the coordinates are
deterministically related.  Then $\E[x\tp{2}] \succ 0$ can easily be
satisfied. 
%, say with $x \sim \normal{0}{I}$.
However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular for
any data distribution, since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1
\cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains components $t
\cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.  Therefore,
Spectral Experts would not be able to identify the parameters of
a mixture of linear regressions for this distribution over $x$.

@@

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
\label{lem:lowRank}
Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\Lop}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
$\|\opX_p^*(\eta_p)\|_\Lop$. The proofs of the following lemmas follow
from standard concentration inequalities, and have been deferred to
\appendixref{sec:proofs}.

%We will appeal to the random design framework that models the input $x$
%as random and show bounds that hold with high probability.
%\paragraph{Adjoint operator}
%In this section, we upper bound the operator norm of the adjoint
%$\|\opX_p(\eta_p)\|_\Lop$.

@@

Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
the data, we will extend the analysis of the robust tensor method in
\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
steps, 2(a) and 2(c) respectively, of
\algorithmref{algo:spectral-experts}.
\begin{lemma}
  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\Lop \le \epsilon$ and
  $\|\hat M_3 - M_3\|_\Lop \le \epsilon$, for some $\epsilon$ such that 
  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\Lop^{-1})\}.$$ 
  Then, there exists a permutation of indices such that  the parameter
  estimates found in step 2 of \algorithmref{algo:spectral-experts}
  satisfy the following with probability at least $1 - \delta$,
  \begin{align*}
  \|\hat \pi - \pi \|_{\infty}
    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\Lop \epsilon ) \\
  \|\hat \beta_h - \beta_h\|_2
    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\Lop^{0.5} \|M_3\|_\Lop \epsilon ).
  \end{align*}
  for all $h \in [k]$.
\end{lemma}

The proof follows by applying standard matrix perturbation results for
the whitening and unwhitening operators, and has again been deferred
\appendixref{sec:proofs}.

@@

\begin{proof}(Theorem 1)
By \reflem{lowRank} along with \reflem{lowRankLower} and
\reflem{lowRankUpper}, we can control the operator norm of the parameter
estimates which are directly upper bounded by the Frobenius norm: If $n
\ge n_1$, then
\begin{align}
  \|\hat M_p - M_p\|_\Lop = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
\end{align}

We complete the proof by applying \reflem{tensorPower} with the above
bound on $\|\hat M_p - M_p\|_\Lop$.

>>>>

While this is not a concern for independent dimensions, we note that
this requires more stringent positive definite constraints when using
features. For example, suppose $x = (1, t, t^2)$, the common polynomial
basis expansion, so that all the coordinates are deterministically
related.  Then $\E[x\tp{2}] \succ 0$ can easily be satisfied. 
%, say with $x \sim \normal{0}{I}$.
However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular for
any data distribution, since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1
\cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains components $t
\cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.  Therefore,
Spectral Experts would not be able to identify the parameters of
a mixture of linear regressions for this distribution over $x$.

@@

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
\label{lem:lowRank}
Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\op}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
$\|\opX_p^*(\eta_p)\|_\op$. The proofs of the following lemmas follow
from standard concentration inequalities and appear in detail in the
supplementary material.
%have been deferred to \appendixref{sec:proofs}.

%We will appeal to the random design framework that models the input $x$
%as random and show bounds that hold with high probability.
%\paragraph{Adjoint operator}
%In this section, we upper bound the operator norm of the adjoint
%$\|\opX_p(\eta_p)\|_\op$.

@@

Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
the data, we will extend the analysis of the robust tensor method in
\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
steps, 2(a) and 2(c) respectively, of
\algorithmref{algo:spectral-experts}.
\begin{lemma}
  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op \le \epsilon$ and
  $\|\hat M_3 - M_3\|_\op \le \epsilon$, for some $\epsilon$ such that 
  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\op^{-1})\}.$$ 
  Then, there exists a permutation of indices such that  the parameter
  estimates found in step 2 of \algorithmref{algo:spectral-experts}
  satisfy the following with probability at least $1 - \delta$,
  \begin{align*}
  \|\hat \pi - \pi \|_{\infty}
    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\op \epsilon ) \\
  \|\hat \beta_h - \beta_h\|_2
    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\op^{0.5} \|M_3\|_\op \epsilon ).
  \end{align*}
  for all $h \in [k]$.
\end{lemma}

The proof follows by applying standard matrix perturbation results for
the whitening and unwhitening operators and has again been deferred
to the supplementary material.
%\appendixref{sec:proofs}.

@@

\begin{proof}[Proof of Theorem 1]
By \reflem{lowRank}, \reflem{lowRankLower} and \reflem{lowRankUpper}, we
can control the Frobenius norm of the error in the moments, which
directly upper bounds the operator norm: If $n \ge \max\{n_1, n_2\}$,
then
\begin{align}
  \|\hat M_p - M_p\|_\op = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
\end{align}

We complete the proof by applying \reflem{tensorPower} with the above
bound on $\|\hat M_p - M_p\|_\op$.
